---
title: "Using DuckDB in R: A Walkthrough Starting with MySQL"
subtitle: ""
excerpt: ""
date: 2024-11-01
author: "Steve Ewing"
draft: false
images:
series:
tags:
categories:
layout: single
---

# Introduction

In this walkthrough, we will demonstrate how to use DuckDB in R, starting by connecting to a MySQL server. By integrating MySQL and DuckDB, you can leverage the strengths of both databases within your data science workflows. This guide will help you retrieve data from MySQL, import it into DuckDB, and perform efficient analytical queries.

## Prerequisites

R and RStudio: Ensure you have R and RStudio installed.
MySQL Server: Access to a MySQL server with the necessary permissions.
Required R Packages: ```RMySQL```, ```DBI```, ```duckdb```, ```dplyr```, ```dbplyr```, ```ggplot2```

## Loading Data into DuckDB From MySQL

First, we need to connect to a MySQL database to retrieve data.

### Installation

```{r}
# install.packages("RMySQL")
# install.packages("DBI")  # Database interface
```

Load the libraries:

```{r}
library(RMySQL)
library(DBI)
```

### Establishing a Connection

Set up the connection to your MySQL server:

```{r}
# Replace with your MySQL server details
con_mysql <- dbConnect(RMySQL::MySQL(),
                       dbname = Sys.getenv("DO_DB_NAME"),
                       host = Sys.getenv("DO_DB_HOST"), 
                       port = as.integer(Sys.getenv("DO_DB_PORT")),
                       user = Sys.getenv("DO_DB_USER"),
                       password = Sys.getenv("DO_DB_PASSWORD"))

```

### Connect to DuckDB

```{r}
# DuckDB connection
con_duckdb <- dbConnect(duckdb::duckdb(), dbdir = "dmo_data.duckdb")
```

### Retrieve the List of Tables from MySQL

Let's retrieve data from a MySQL table:

```{r}
# List available tables
mysql_tables <- dbListTables(con_mysql)

# View the list of tables
print(mysql_tables)
```

If you want to load only specific tables, you can define them:

```{r}
# Specify the tables you want to load
tables_to_load <- c("orders", "products", "orders_products", "orders_products_pick", "orders_total", "shipments", "ups_import", "products_to_products_groups", "products_groups", "fees", "payments_charges", "payments_refunds", "payments_types", "wm_po_items", "wm_purchase_orders", "wm_boxes", "autoship_orders", "box_subscription_orders", "products_bundles", "products_inventory", "wm_inventory_transactions", "wm_suppliers", "wm_invoices", "wm_po_history", "shipments_status", "orders_status", "manufacturers", "customers", "customers_credit_accounts", "address_book", "wm_product_promos", "wm_product_promos_products", "wm_product_promos_tiers", "customers_groups", "suppliers_import", "wm_suppliers_to_products", "products_to_products_extra_fields", "products_extra_fields")
```

### Transfer Tables from MySQL to DuckDB

We'll iterate over each table, read it from MySQL, and write it into DuckDB.

```{r}
# Use the full list or your specified list
tables <- tables_to_load  # or use 'mysql_tables' for all tables

for (table_name in tables) {
  # Read data from MySQL
  message("Loading table: ", table_name)
  data <- dbReadTable(con_mysql, table_name)
  
  # Write data to DuckDB
  dbWriteTable(con_duckdb, table_name, data, overwrite = TRUE)
}
```

### Disconnect From the Server

```{r}
# Disconnect from MySQL
dbDisconnect(con_mysql)
```

### Verify the Tables in DuckDB

List the tables in DuckDB to confirm they have been loaded:

```{r}
duckdb_tables <- dbListTables(con_duckdb)
print(duckdb_tables)
```

## Performing Queries in DuckDB

With the tables loaded into DuckDB, you can perform efficient queries using SQL or dplyr.

### Using SQL Queries

```{r}
result <- dbGetQuery(con_duckdb, "
  SELECT op.products_id, 
         COUNT(op.qty_returned) AS returns
  FROM orders_products op
  LEFT JOIN products p USING(products_id)
  LEFT JOIN orders o USING(orders_id)
  WHERE op.qty_returned > 0
    AND op.qty_returned IS NOT NULL
  GROUP BY op.products_id, p.name
  ORDER BY returns DESC
  LIMIT 10;
")

print(result)

```
